### 2024-12-05 manuscript.tex
Addressing comments

The main aim of this study is to evaluate the GSRM version of ICON, developed
jointly by nextGEMS, Deutscher Wetterdienst, Max-Planck-Institute for
Meteorology, Deutsches Klimarechenzentrum (DKRZ), Karlsruhe Institute of
Technology, and the Center for Climate Systems Modeling.  Previous studies have
identified substantial large-scale biases in climate model clouds over the SO,
affecting sea surface temperature and the Earth’s albedo. Our aim is to
quantify how well the GSRM ICON is simulating clouds in this region,
particularly in light of the fact that subgrid-scale clouds and convection are
not parameterized in this model. This region is mostly dominated by boundary
layer clouds generated by shallow convection, and these are problematic to
observe by spaceborne lidars and radars, which are affected by attenuation by
overlapping and thick clouds and ground clutter, respectively.  Specifically,
the radar on CloudSat and lidar on CALIPSO (neither of which are nowoperational) are affected by the abovementioned issues, resulting in a strong
underestimation of cloud occurrence below 2 km relative to ground-based lidar
observations \citep{mcerlich2021}.  We hypothesise that this, in turn, can lead to systematic
biases in low clouds in climate models, which are frequently evaluated against
CloudSat--CALIPSO products. Reanalyses can also suffer from cloud biases, as
these are usually parametrised in their atmospheric component, and also in
regions where input observations are sparse.  This makes them a problematic
reference for clouds over the SO, and any biases relative to a reanalysis
should be interpreted with caution. Instead, we chose to use a large set of
ship-based observations conducted with ceilometers and lidars on board of the
RV \emph{Polarstern} and other voyages and stations as a reference for the
model evaluation.

@@

Unlike current general circulation models (GCMs), the storm-resolving version
of ICON does not use convective and cloud parameterization but relies on
explicit simulation of convection and clouds on the model grid. While this
makes the code development simpler without having to rely on uncertain
parameterizations, it can miss smaller-scale clouds below the grid resolution.
Turbulence and cloud microphysics are still parameterized in this model. To
account for the radiative effects of subgrid-scale clouds, a cloud
inhomogeneity factor is introduced in the model which scales down the cloud
liquid water for radiative calculations. It ranges from 0.4 at lower
tropospheric stability (LTS) of 0 K to 0.8 at 30 K. In addition, vertical
mixing is enhanced in unstable and humid lower-tropospheric conditions, which
reduces the amount of shallow clouds.

@@

\begin{figure}[p!]
\centerline{
\includegraphics[width=1.06\textwidth]{img/cloud_occurrence_panel.pdf}
}
\caption{Cloud occurrence by height for the 31 voyages and one sub-Antarctic
station (MICRE) in observations (OBS) and simulated by the ALCF from the
ICON model, MERRA‐2 (M2), and ERA5 reanalysis data. The numbers in the legend
indicate the total cloud fraction and the number of days of data.}
\label{fig:cloud-occurrence-panel}
\end{figure}

>>>>

The main aim of this study is to evaluate the GSRM version of ICON. ICON developed and maintained
jointly by Deutscher Wetterdienst, Max-Planck-Institute for
Meteorology, Deutsches Klimarechenzentrum (DKRZ), Karlsruhe Institute of
Technology, and the Center for Climate Systems Modeling.  Previous studies have
identified substantial large-scale biases in climate model clouds over the SO,
affecting sea surface temperature and the Earth’s albedo. Our aim is to
quantify how well the GSRM ICON is simulating clouds in this region,
particularly in light of the fact that subgrid-scale clouds and convection are
not parameterized in this model. This region is mostly dominated by boundary
layer clouds generated by shallow convection, and these are problematic to
observe by spaceborne lidars and radars, which are affected by attenuation by
overlapping and thick clouds and ground clutter, respectively.  Specifically,
the radar on CloudSat and lidar on CALIPSO (neither of which are nowoperational) are affected by the abovementioned issues, resulting in a strong
underestimation of cloud occurrence below 2 km relative to ground-based lidar
observations \citep{mcerlich2021}.  We hypothesise that this, in turn, can lead to systematic
biases in low clouds in climate models, which are frequently evaluated against
CloudSat--CALIPSO products. Reanalyses can also suffer from cloud biases, as
these are usually parametrised in their atmospheric component, and also in
regions where input observations are sparse.  This makes them a problematic
reference for clouds over the SO, and any biases relative to a reanalysis
should be interpreted with caution. Instead, we chose to use a large set of
ship-based observations conducted with ceilometers and lidars on board of the
RV \emph{Polarstern} and other voyages and stations as a reference for the
model evaluation.

@@

Unlike current general circulation models (GCMs), the storm-resolving version
of ICON does not use convective and cloud parameterization but relies on
explicit simulation of convection and clouds on the model grid. Subgrid-scale
clouds are not resolved, and the grid cell cloud fraction is always either 0 or
100\%. While this makes the code development simpler without having to rely on
uncertain parameterizations, it can miss smaller-scale clouds below the grid
resolution.  Turbulence and cloud microphysics are still parameterized in this
model. To account for the radiative effects of subgrid-scale clouds, a cloud
inhomogeneity factor is introduced in the model which scales down the cloud
liquid water for radiative calculations. It ranges from 0.4 at lower
tropospheric stability (LTS) of 0 K to 0.8 at 30 K. In addition, vertical
mixing is enhanced in unstable and humid lower-tropospheric conditions, which
reduces the amount of shallow clouds.

@@

\begin{figure}[p!]
\centerline{
\includegraphics[width=1.06\textwidth]{img/cloud_occurrence_panel.pdf}
}
\caption{
Cloud occurrence by height for the 31 voyages and one sub-Antarctic
station (MICRE) in observations (O) and simulated by the ALCF from the ICON
model (I), MERRA‐2 (M), and ERA5 reanalysis data (E). The numbers in the legend
indicate the total cloud fraction and the number of days of data.
}
\label{fig:cloud-occurrence-panel}
\end{figure}
